---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r, echo = TRUE}
unzip(zipfile = "activity.zip", exdir = "data")
activity <- read.csv(file = "data/activity.csv")
```

## What is mean total number of steps taken per day?

### 1. Calculating total number of steps per day:
```{r, echo = TRUE}
stepsByDate <- aggregate(steps ~ date, activity, FUN = sum)
```

### 2. Making histogram of total number of steps per each day
```{r, echo = TRUE}
library(ggplot2)
g <- ggplot(data = stepsByDate, aes(steps))
p <- g + geom_histogram(binwidth=1000) +
         labs(title = "Histogram of total number of steps taken each day") + 
         xlab("Total numbers of steps in day")
p
```

**NOTE** According to `binwidth` parameter histogram can looks like different. 
I decided to use binwidth = 1000 due to diagram seems to be more readable.

### 3. Calculating mean and median 
```{r, echo = TRUE}
mean(stepsByDate$steps)
median(stepsByDate$steps)
```

## What is the average daily activity pattern?

### 1. Creating time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis):
```{r, echo = TRUE}
meanStepsByInterval <- aggregate(steps ~ interval, activity, FUN = mean)

g <- ggplot(data = meanStepsByInterval, aes(x = interval, y = steps))
p <- g + geom_line() +
         labs(title = "Average number of steps by 5-minute interval") + 
         xlab("time interval")
p

```

### 2. Finding 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps:
```{r, echo = TRUE}
meanStepsByInterval[which.max(meanStepsByInterval$steps),1:2]
```

## Imputing missing values

### 1. Calculating the total number of missing values in the dataset:
```{r, echo = TRUE}
length(activity$steps[is.na(activity$steps)])

```

### 2. Strategy for missing vlue imputing.
Missing values will be replaced by the mean value of particular 5-minute interval. 
These values already calculated in `meanStepsByInterval$steps`.

### 3. Creating new dataset with filled steps missing values (`actFilled`)
```{r, echo = TRUE}
actFilled <- cbind(activity, meanStepsByInterval$steps)
names(actFilled)[4]<-"meanStepsByInterval"
na_steps <- is.na(actFilled$steps)
actFilled$steps[na_steps] <- actFilled$meanStepsByInterval[na_steps]

```

### 4. Make histogram for new dataset and calculate mean and median.

```{r, echo = TRUE}
stepsByDateFilled <- aggregate(steps ~ date, actFilled, FUN = sum)

g <- ggplot(data = stepsByDateFilled, aes(steps))
p <- g + geom_histogram(binwidth=1000) +
         labs(title = "Histogram of total number of steps taken each day") + 
         xlab("Total numbers of steps in day")
p

```


Mean and median for dataset without NAs:
```{r, echo = TRUE}
mean(stepsByDateFilled$steps)
median(stepsByDateFilled$steps)
```
Mean and median for original dataset (with NAs):
```{r, echo = TRUE}
mean(stepsByDate$steps)
median(stepsByDate$steps)
```

**As a result of missing values imputing `mean` value was not changed. But `median` value was shightly changed.**

## Are there differences in activity patterns between weekdays and weekends?

### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend”

```{r, echo = TRUE}
# set locale to English to print weekday names in English.
Sys.setlocale("LC_ALL","English")
actFilled$date <- as.Date(actFilled$date)
actFilled$weekDayType <- "weekday"
actFilled$weekDayType[weekdays(actFilled$date) %in% c("Sunday", "Saturday")] <- "weekend"
actFilled$weekDayType <- as.factor(actFilled$weekDayType)
```

### 2. Make plot of time series of 5-minute interval and the avg number of steps across weekdays or weekends

```{r, echo = TRUE}
meanStepsByIntervalFilled <- aggregate(steps ~ interval+weekDayType, actFilled, FUN = mean)

g <- ggplot(data = meanStepsByIntervalFilled, aes(x = interval, y = steps))
p <- g + geom_line() +
         facet_wrap(facets = ~ weekDayType, nrow = 2, ncol = 1) +
         labs(title = "Average number of steps for 5-minute interval by day type") + 
         xlab("time interval")
p

```


